---
title: "Plotly"
author: "Rajat Jain"
output:
  ioslides_presentation:
    widescreen: true
    fig_caption: true
  pdf_document: 
    always_allow_html: yes
  html_document:
    theme: lumen
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r libs, message=FALSE, warning=FALSE, include=FALSE, results='hide'}
library(magrittr)
library(wordcloud)
library(ggplot2)
library(dlstats)
library(plotly)

rand <- function() {
  return ((sample(1:100)[1]) / 100.0)
}
```


## Plotly's R graphing library
Makes interactive, publication-quality Graphs.
```{r wordcloud, fig.align="center", fig.height=4.5, fig.width=9, message=FALSE, warning=FALSE, echo=FALSE}
wordcloud(c("line plots", "scatter plots", "area charts", "bar charts", "error bars", "box plots", "histograms", "heatmaps", "subplots", "multiple-axes", "3D WebGL based charts"), min.freq=1, max.words = 100, random.order = TRUE, random.color = TRUE, colors = rev(colorRampPalette(brewer.pal(9,"Blues"))(32)[seq(8,32,6)]))
```
Courtesy: https://plot.ly/r/


## Data
For this presentation we will use package download statistics data from CRAN.
```{r data, fig.align="center"}
pkg.downloads <- cran_stats(c("ggplot2", "ggvis", "plotly", "dygraphs", "ggiraph"))
head(pkg.downloads)
```
Data Source: dlstats package - https://cran.r-project.org/web/packages/dlstats/index.html


## A Simple ggplot2 line chart
```{r ggplot, fig.align="center", fig.height=4, fig.width=9, message=FALSE, warning=FALSE}
simple_plot <- ggplot(data=pkg.downloads, aes(x=start, y=downloads, color=package)) +
  geom_line() + xlab("Months") + ylab("# Downloads") +
  labs(title="Package Adoption", color="Package")
simple_plot
```


## Plotly in Action
```{r plotly, fig.align="center", fig.height=4, fig.width=9, message=FALSE, warning=FALSE}
plot_ly(data=pkg.downloads, x=~start, y=~downloads, color=~package, mode='lines') %>%
  layout(title="Package Adoption", 
         xaxis=list(title="Months"), yaxis=list(title="# Downloads"))
```


## What if we have multiple statistics to plot?
We have only one statistic - downloads in the original dataset. Let's create another one - installs. Successful installs are a certain percentage (<= 100%) of downloads.
```{r installs, fig.align="center"}
pkg.downloads$installs = round(rand() * pkg.downloads$downloads)
head(pkg.downloads)
```


## Plotly - Multiple Traces
```{r multitrace, fig.align="center", fig.height=4, fig.width=9, message=FALSE, warning=FALSE}
mt_plot <- plot_ly(data=pkg.downloads[pkg.downloads$package=='plotly',], x=~start) %>%
  add_trace(y=~downloads, name='Downloads', mode='lines') %>%
  add_trace(y=~installs, name='Installs', mode='lines+markers') %>%
  layout(title="Plotly Download & Install Success") 
mt_plot
```


## Like It? Share It!
```{r auth, message=FALSE, warning=FALSE, include=FALSE, results='hide', echo=FALSE}
Sys.setenv("plotly_username"="rajatj")
Sys.setenv("plotly_api_key"="")
```
Create a shareable link to your chart using Plotly API.
```{r share, message=FALSE, warning=FALSE}
chart_link <- api_create(mt_plot, filename="multi-trace-plot1")
```
Saved at URL: `r chart_link$web_url`

Before using this feature, you will have to setup API credentials. Follow instructions here: https://plot.ly/r/getting-started/#initialization-for-online-plotting


## Got ggplot2 Charts?
No worries, Plotly can transform them!
```{r transform, fig.align="center", fig.height=4, fig.width=9, message=FALSE, warning=FALSE}
ggplotly(simple_plot)
```


## Similar
There are several Graphing packages available for R. E.g.,

- ggplot2
- ggvis

Choice is yours,

* ggplot2 continues to be the most popular one. We have proof!
* plotly is great with interactive charts
    + Good for dashboards
    + Not so much, if your final output is non-HTML like PDF - https://plot.ly/r/knitr/#embedding-plotly-graphs-in-rmd-files

Ref: Nice blog comparing ggplot2 and plotly - https://www.bryanwhiting.com/2017/02/the-dataviz-battle-plotly-vs-ggplot2/


## References

- Getting Started: https://plot.ly/r/getting-started/

- Everything Plotly: https://plot.ly/r/reference/

- Surprisingly good Cheet Sheet: https://images.plot.ly/plotly-documentation/images/r_cheat_sheet.pdf

- R Markdown file which generated this presentation: https://github.com/rajatmnnit/csx415-assignments/tree/master/00-package-presentations/plotly/plotly.Rmd